\bf Title here

### 2020-07-21 paper/paper.Rmd
some re-writing of exploring index value and update plots

## Visual diagnostics

@@

### Explore the value of objective function {#static}

It is a convention to map the time index of a time series on the x-axis and value to the y-axis. In our diagnostic plots, it is to map the iterator on the x-axis and index value to the y-axis. The next question is to ask, given four indices being defined in the global object, which one should be used? *It depends on the *.

As the searching goes towards the end, it is easy to have hundreds of candidate bases in each iteration with very close index value. When plotting, these iteration will occupy the vast majority of the plot space. This motivates to summarise the points in each iteration (`tries`). At each iteration, rather than knowing the index value of *every* points, we are more interested to have a general summary of all the points and more importantly, the point with the largest `index_val` since it prescribes the next geodesic interpolation and future searches.

Boxplot is a suitable candidate that provides five points summary of the data, however it has one drawback: it doesn't report the number of point in each box. We may risk losing information on how many points it takes to find the target basis by displaying the boxplot alone for all `tries`. Thus, the number of point for each `tries` is displayed at the bottom of each box and we provide options to switch `tries` with small number of points to a point geometry, which is achieved via the `cutoff` argument. A line geometry is also added to link the points with the largest index value in each `tries`. This helps to visualise the improvement made by each `tries`. With gramma of graphics [@wickham2010layered], the plot for exploring index value can be defined in three layers as follows:

@@

The larger dataset is used for 2D projection with `search_better` and `max.tries = 500`. In Figure \ref{points}, the searching points with `id` and `tries` on the x-axis, colored by `tries`. Label at the bottom indicates the number of observations in each tries and facilitates the choice of cutoff to switch from point geometry to boxplot geometry (`cutoff = 15`). The line geometry suggests the largest improvement happens at `tries = 5`.

@@

Sometimes, it is more interesting to explore the index value on the interpolation path and the plot definition is slightly varied.

@@

This examples uses `search_better` for a 2D projection on the larger dataset using the `holes` index. In the interpolation stage, continuous frames will be constructed to bridge the project on the current basis to the target basis.  The target basis will become the current basis in the next iteration and the searching stage continues to find the next target basis. In the left panel of Figure \ref{interruption}, we observe that  there are interpolating bases with higher index values than the target basis and these bases could be used to search for new basis in the next searching stage.

An interruption is then constructed to accept the interpolating bases up to the one with the largest index value and use that basis as the current basis for the next searching stage. After implementing this interruption, the tracing plot with the same configuration is shown on the right panel of Figure \ref{interruption}. In the third `tries`, rather than interpolating fully to the target basis at `id = 62`, it stops before the index value starts to decrease at `id = 60`. This implementation results in a higher index value in the end with fewer steps.

@@

@@

In principle, all the optimisation routines should result in the same output for the same problem while this may not be true in real application. This motivates the creation of a polishing search that polishes the final basis and achieves unity on different methods.

`search_polish` takes the final basis of a given search as the current basis and uses a brutal-force approach to sample a large number of basis (`n_sample`) in the neighbourhood, whose radius is controlled by `polish_alpha`. Among the `n_sample` basis, the one with the largest index value becomes the candidate basis. If its index value of the candidate basis is larger than that of the current basis, it becomes the current basis in the next iteration. If no basis is found to have larger index value than the current basis, the searching neighbourhood will be shrunk and the search continues. The polishing search ends when one of the four stopping criteria is satisfied:

1) the two basis can't be too close 
2) the percentage improvement of the index function can't be too small 
3) the searching neighbourhood can't be too small
4) the number of iteration can't exceed the `max.tries`

[should the stopping criteria be more detailed? xxx]

The usage of search_polish is as follows. After the first tour, the final basis from the interpolation is extracted and supplied into a new tour with the `start` argument and `search_polish` as the searching function in the guided_tour. All the other arguments should remain the same.

@@

The following example conducted a 2D projection on the larger dataset using search better with different configurations. `max.tries` is a hyperparameter that controls the maximum number of try without improvement and its default value is 25. As shown in Figure \ref{trace-compare}, both trials attain the same index value after polishing while a small `max.tries` of 25 is not sufficient for `search_better` to find the global maximum. The `max.tries` argument needs to be adjusted to ensure it is sufficient to explore the parameter space.

@@

### Explore searching space {#animated}

@@

#### A toy example: understand different stage of search_geodesic

@@

#### A more complex example: Choosing the initial value for polishing parameter

@@

@@

### A comprehensive example of diagnosing a noisy index function

@@

#### Explore index value

@@

#### Explore searching space

>>>>

# Visual diagnostics

@@

## Explore the value of objective function {#static}

### Searching points

A primary interest of diagnosing an optimisation algorithm is to study how it finds its optimal progressively. We could plot the index value across its natural ordering, however, different iterations may have different number of points and, towards the end of the search there could easily be hundreds of bases being tested before the target basis is found. In the plot, points from those iterations towards the end will occupy the vast majority of the plot space. This motivates to use summarisation.  Rather than knowing the index value of *every* basis, we are more interested to have a general summary of all the index value in that iteration and more importantly, the basis with the largest index value (since it prescribes the next geodesic interpolation and future searches).

Boxplot is a suitable candidate that provides five points summary of the data, however it has one drawback: it doesn't report the number of point in each box. We may risk losing information on how many points it takes to find the target basis by displaying the boxplot alone for all `tries`. Thus, the number of point in each iteration is displayed at the bottom of each box and we provide options to switch iteration with small number of points to a point geometry, which is achieved via an `cutoff` argument. A line geometry is also added to link the points with the largest index value in each iteration. This helps to visualise the improvement made in each iteration. Using the concept of *gramma of graphics* [@wickham2010layered], the plot for exploring index value can be defined in three layers as:

@@

We choose variable `x1`, `x2`, `x3`, `x8`, `x9` and `x10` to perform a 2D projection with tour. Parameter `search_f = tour::search_better` and `max.tries = 500` is used. The index value of the searching points are shown in Figure \ref{points}. Label at the bottom indicates the number of observations in each iteration and facilitates the choice of `cutoff` argument (by default `cutoff = 15`). We learn that the `search_better` quickly finds better projection basis with higher index value at first and the takes longer to find a better one later.

@@

### Interpolating points

Sometimes, rather than explore the searching points, we may be interested in explore the points on the interpolation path (target and interpolating points) since these points will be played by the tour animation. Since interpolation paths are geodesically the shortest, a summarisation using boxplot geometry is no longer needed. The slightly modified plot definition is shown below:

@@

We use the same dataset as the toy example above to explore the search function `search_better` and we want to learn how the index value changes on the interpolation path for `holes` index. From the left panel of Figure \ref{interruption}, we observe that when interpolating from the current basis to the target basis, the index value may not be monotone: we could reach a basis with higher index value than the target basis on the interpolation path. In this sense, we would be better off using the basis with the highest index value on the interpolation path as the current basis for the next iteration (rather than using the target basis). Hence, an interruption is constructed to accept the interpolating bases only up to the one with the largest index value. After implementing this interruption, the search finds higher final index value with fewer steps as shown in the right panel of Figure \ref{interruption}.

@@

### Polishing points

@@

In principle, all the optimisation routines should result in the same output for the same problem while this may not be the case in real application. This motivates the creation of a polishing search that polishes the final basis found and achieves unity across different methods.

`search_polish` takes the final basis of a given search as a start and uses a brutal-force approach to sample a large number of basis (`n_sample`) in the neighbourhood. Among those sampled basis, the one with the largest index value is chosen to be compared with the current basis. If its index value is larger than that of the current basis, it becomes the current basis in the next iteration. If no basis is found to have larger index value, the searching neighbourhood will shrink and the search continues. The polishing search ends when one of the four stopping criteria is satisfied:

1) the chosen basis can't be too close to the current basis 
2) the percentage improvement of the index value can't be too small 
3) the searching neighbourhood can't be too small
4) the number of iteration can't exceed the `max.tries`

The usage of search_polish is as follows. After the first search, the final basis from the interpolation is extracted and supplied to the second search as the `start` argument.  `search_polish` is used as the search function. All the other arguments should remain the same.

@@

Again using the same data, we are interested to compare the effect of different `max.tries` in the 2D projection setting. `max.tries` is a hyperparameter that controls the maximum number of try before the search ends. The default value of 25 is suitable for 1D projection while we suspect it may not be a good option for the 2D case and hence want to compare it with an alternative, 500. As shown in Figure \ref{trace-compare}, both trials attain the same index value after polishing while the small `max.tries` of 25 is not sufficient for `search_better` to find its global maximum and we will need to adjust the `max.tries` argument for the search to succiciently explore the parameter space.

@@

## Explore searching space {#animated}

@@

### A toy example: understand different stage of search_geodesic

@@

### A more complex example: Choosing the initial value for polishing parameter

@@

[add example from geozoo]

@@

## A comprehensive example of diagnosing a noisy index function

@@

### Explore index value

@@

### Explore searching space

### 2020-07-21 paper/paper.tex
some re-writing of exploring index value and update plots

\hypertarget{visual-diagnostics}{%
\subsection{Visual diagnostics}\label{visual-diagnostics}}

@@

\hypertarget{static}{%
\subsubsection{Explore the value of objective function}\label{static}}

It is a convention to map the time index of a time series on the x-axis
and value to the y-axis. In our diagnostic plots, it is to map the
iterator on the x-axis and index value to the y-axis. The next question
is to ask, given four indices being defined in the global object, which
one should be used? \emph{It depends on the }.

As the searching goes towards the end, it is easy to have hundreds of
candidate bases in each iteration with very close index value. When
plotting, these iteration will occupy the vast majority of the plot
space. This motivates to summarise the points in each iteration
(\texttt{tries}). At each iteration, rather than knowing the index value
of \emph{every} points, we are more interested to have a general summary
of all the points and more importantly, the point with the largest
\texttt{index\_val} since it prescribes the next geodesic interpolation
and future searches.

Boxplot is a suitable candidate that provides five points summary of the
data, however it has one drawback: it doesn't report the number of point
in each box. We may risk losing information on how many points it takes
to find the target basis by displaying the boxplot alone for all
\texttt{tries}. Thus, the number of point for each \texttt{tries} is
displayed at the bottom of each box and we provide options to switch
\texttt{tries} with small number of points to a point geometry, which is
achieved via the \texttt{cutoff} argument. A line geometry is also added
to link the points with the largest index value in each \texttt{tries}.
This helps to visualise the improvement made by each \texttt{tries}.
With gramma of graphics \citep{wickham2010layered}, the plot for
exploring index value can be defined in three layers as follows:

@@

The larger dataset is used for 2D projection with
\texttt{search\_better} and \texttt{max.tries\ =\ 500}. In Figure
\ref{points}, the searching points with \texttt{id} and \texttt{tries}
on the x-axis, colored by \texttt{tries}. Label at the bottom indicates
the number of observations in each tries and facilitates the choice of
cutoff to switch from point geometry to boxplot geometry
(\texttt{cutoff\ =\ 15}). The line geometry suggests the largest
improvement happens at \texttt{tries\ =\ 5}.

@@

Sometimes, it is more interesting to explore the index value on the
interpolation path and the plot definition is slightly varied.

@@

This examples uses \texttt{search\_better} for a 2D projection on the
larger dataset using the \texttt{holes} index. In the interpolation
stage, continuous frames will be constructed to bridge the project on
the current basis to the target basis. The target basis will become the
current basis in the next iteration and the searching stage continues to
find the next target basis. In the left panel of Figure
\ref{interruption}, we observe that there are interpolating bases with
higher index values than the target basis and these bases could be used
to search for new basis in the next searching stage.

An interruption is then constructed to accept the interpolating bases up
to the one with the largest index value and use that basis as the
current basis for the next searching stage. After implementing this
interruption, the tracing plot with the same configuration is shown on
the right panel of Figure \ref{interruption}. In the third
\texttt{tries}, rather than interpolating fully to the target basis at
\texttt{id\ =\ 62}, it stops before the index value starts to decrease
at \texttt{id\ =\ 60}. This implementation results in a higher index
value in the end with fewer steps.

@@

In principle, all the optimisation routines should result in the same
output for the same problem while this may not be true in real
application. This motivates the creation of a polishing search that
polishes the final basis and achieves unity on different methods.

\texttt{search\_polish} takes the final basis of a given search as the
current basis and uses a brutal-force approach to sample a large number
of basis (\texttt{n\_sample}) in the neighbourhood, whose radius is
controlled by \texttt{polish\_alpha}. Among the \texttt{n\_sample}
basis, the one with the largest index value becomes the candidate basis.
If its index value of the candidate basis is larger than that of the
current basis, it becomes the current basis in the next iteration. If no
basis is found to have larger index value than the current basis, the
searching neighbourhood will be shrunk and the search continues. The
polishing search ends when one of the four stopping criteria is
satisfied:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  the two basis can't be too close
\item
  the percentage improvement of the index function can't be too small
\item
  the searching neighbourhood can't be too small
\item
  the number of iteration can't exceed the \texttt{max.tries}
\end{enumerate}

{[}should the stopping criteria be more detailed? xxx{]}

The usage of search\_polish is as follows. After the first tour, the
final basis from the interpolation is extracted and supplied into a new
tour with the \texttt{start} argument and \texttt{search\_polish} as the
searching function in the guided\_tour. All the other arguments should
remain the same.

@@

The following example conducted a 2D projection on the larger dataset
using search better with different configurations. \texttt{max.tries} is
a hyperparameter that controls the maximum number of try without
improvement and its default value is 25. As shown in Figure
\ref{trace-compare}, both trials attain the same index value after
polishing while a small \texttt{max.tries} of 25 is not sufficient for
\texttt{search\_better} to find the global maximum. The
\texttt{max.tries} argument needs to be adjusted to ensure it is
sufficient to explore the parameter space.

@@

\hypertarget{animated}{%
\subsubsection{Explore searching space}\label{animated}}

@@

\hypertarget{a-toy-example-understand-different-stage-of-search_geodesic}{%
\paragraph{A toy example: understand different stage of
search\_geodesic}\label{a-toy-example-understand-different-stage-of-search_geodesic}}

@@

\hypertarget{a-more-complex-example-choosing-the-initial-value-for-polishing-parameter}{%
\paragraph{A more complex example: Choosing the initial value for
polishing
parameter}\label{a-more-complex-example-choosing-the-initial-value-for-polishing-parameter}}

@@

\hypertarget{a-comprehensive-example-of-diagnosing-a-noisy-index-function}{%
\subsubsection{A comprehensive example of diagnosing a noisy index
function}\label{a-comprehensive-example-of-diagnosing-a-noisy-index-function}}

@@

\hypertarget{explore-index-value}{%
\paragraph{Explore index value}\label{explore-index-value}}

@@

\hypertarget{explore-searching-space}{%
\paragraph{Explore searching space}\label{explore-searching-space}}

>>>>

\hypertarget{visual-diagnostics}{%
\section{Visual diagnostics}\label{visual-diagnostics}}

@@

\hypertarget{static}{%
\subsection{Explore the value of objective function}\label{static}}

\hypertarget{searching-points}{%
\subsubsection{Searching points}\label{searching-points}}

A primary interest of diagnosing an optimisation algorithm is to study
how it finds its optimal progressively. We could plot the index value
across its natural ordering, however, different iterations may have
different number of points and, towards the end of the search there
could easily be hundreds of bases being tested before the target basis
is found. In the plot, points from those iterations towards the end will
occupy the vast majority of the plot space. This motivates to use
summarisation. Rather than knowing the index value of \emph{every}
basis, we are more interested to have a general summary of all the index
value in that iteration and more importantly, the basis with the largest
index value (since it prescribes the next geodesic interpolation and
future searches).

Boxplot is a suitable candidate that provides five points summary of the
data, however it has one drawback: it doesn't report the number of point
in each box. We may risk losing information on how many points it takes
to find the target basis by displaying the boxplot alone for all
\texttt{tries}. Thus, the number of point in each iteration is displayed
at the bottom of each box and we provide options to switch iteration
with small number of points to a point geometry, which is achieved via
an \texttt{cutoff} argument. A line geometry is also added to link the
points with the largest index value in each iteration. This helps to
visualise the improvement made in each iteration. Using the concept of
\emph{gramma of graphics} \citep{wickham2010layered}, the plot for
exploring index value can be defined in three layers as:

@@

We choose variable \texttt{x1}, \texttt{x2}, \texttt{x3}, \texttt{x8},
\texttt{x9} and \texttt{x10} to perform a 2D projection with tour.
Parameter \texttt{search\_f\ =\ tour::search\_better} and
\texttt{max.tries\ =\ 500} is used. The index value of the searching
points are shown in Figure \ref{points}. Label at the bottom indicates
the number of observations in each iteration and facilitates the choice
of \texttt{cutoff} argument (by default \texttt{cutoff\ =\ 15}). We
learn that the \texttt{search\_better} quickly finds better projection
basis with higher index value at first and the takes longer to find a
better one later.

@@

\hypertarget{interpolating-points}{%
\subsubsection{Interpolating points}\label{interpolating-points}}

Sometimes, rather than explore the searching points, we may be
interested in explore the points on the interpolation path (target and
interpolating points) since these points will be played by the tour
animation. Since interpolation paths are geodesically the shortest, a
summarisation using boxplot geometry is no longer needed. The slightly
modified plot definition is shown below:

@@

We use the same dataset as the toy example above to explore the search
function \texttt{search\_better} and we want to learn how the index
value changes on the interpolation path for \texttt{holes} index. From
the left panel of Figure \ref{interruption}, we observe that when
interpolating from the current basis to the target basis, the index
value may not be monotone: we could reach a basis with higher index
value than the target basis on the interpolation path. In this sense, we
would be better off using the basis with the highest index value on the
interpolation path as the current basis for the next iteration (rather
than using the target basis).

Hence, an interruption is constructed to accept the interpolating bases
only up to the one with the largest index value. After implementing this
interruption, the search finds higher final index value with fewer steps
as shown in the right panel of Figure \ref{interruption}.

@@

\hypertarget{polishing-points}{%
\subsubsection{Polishing points}\label{polishing-points}}

In principle, all the optimisation routines should result in the same
output for the same problem while this may not be the case in real
application. This motivates the creation of a polishing search that
polishes the final basis found and achieves unity across different
methods.

\texttt{search\_polish} takes the final basis of a given search as a
start and uses a brutal-force approach to sample a large number of basis
(\texttt{n\_sample}) in the neighbourhood. Among those sampled basis,
the one with the largest index value is chosen to be compared with the
current basis. If its index value is larger than that of the current
basis, it becomes the current basis in the next iteration. If no basis
is found to have larger index value, the searching neighbourhood will
shrink and the search continues. The polishing search ends when one of
the four stopping criteria is satisfied:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  the chosen basis can't be too close to the current basis
\item
  the percentage improvement of the index value can't be too small
\item
  the searching neighbourhood can't be too small
\item
  the number of iteration can't exceed the \texttt{max.tries}
\end{enumerate}

The usage of search\_polish is as follows. After the first search, the
final basis from the interpolation is extracted and supplied to the
second search as the \texttt{start} argument. \texttt{search\_polish} is
used as the search function. All the other arguments should remain the
same.

@@

Again using the same data, we are interested to compare the effect of
different \texttt{max.tries} in the 2D projection setting.
\texttt{max.tries} is a hyperparameter that controls the maximum number
of try before the search ends. The default value of 25 is suitable for
1D projection while we suspect it may not be a good option for the 2D
case and hence want to compare it with an alternative, 500. As shown in
Figure \ref{trace-compare}, both trials attain the same index value
after polishing while the small \texttt{max.tries} of 25 is not
sufficient for \texttt{search\_better} to find its global maximum and we
will need to adjust the \texttt{max.tries} argument for the search to
succiciently explore the parameter space.

@@

\hypertarget{animated}{%
\subsection{Explore searching space}\label{animated}}

@@

\hypertarget{a-toy-example-understand-different-stage-of-search_geodesic}{%
\subsubsection{A toy example: understand different stage of
search\_geodesic}\label{a-toy-example-understand-different-stage-of-search_geodesic}}

@@

\hypertarget{a-more-complex-example-choosing-the-initial-value-for-polishing-parameter}{%
\subsubsection{A more complex example: Choosing the initial value for
polishing
parameter}\label{a-more-complex-example-choosing-the-initial-value-for-polishing-parameter}}

@@

\hypertarget{a-comprehensive-example-of-diagnosing-a-noisy-index-function}{%
\subsection{A comprehensive example of diagnosing a noisy index
function}\label{a-comprehensive-example-of-diagnosing-a-noisy-index-function}}

@@

\hypertarget{explore-index-value}{%
\subsubsection{Explore index value}\label{explore-index-value}}

@@

\hypertarget{explore-searching-space}{%
\subsubsection{Explore searching space}\label{explore-searching-space}}
